\texttt{MSnbase}, efficient and elegant R-based processing and visualisation of raw mass spectrometry data

### 2020-07-24 msnbase2.Rnw
triplicated reading timings

<<time_sz>>=
load("bench_time_sz.rda")
time <- sapply(time_sz, function(x) x[, "time"])
sz <- sapply(time_sz, function(x) x[, "sz"])
colnames(sz) <- colnames(time) <- c(1, 5, 10)
rownames(sz) <- rownames(time) <- c("in-memory", "on-disk")
sz <- round(sz/(1014^2), 2)

p_read <-
    data.frame(time) %>%
    rownames_to_column() %>%
    pivot_longer(names_to = "files",
                 values_to = "time",
                 -rowname) %>%
    rename(backend = rowname) %>%
    mutate(files = as.numeric(sub("X", "", files))) %>%
    ggplot(aes(x = files, y = time, colour = backend)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks=c(1, 5, 10)) +
    ylab("Reading time [s]") +
    xlab("Number of files") +
    theme(legend.position = c(0.35, 0.86),
          legend.title = element_blank(),
          legend.text=element_text(size = 6))

p_sz <-
    data.frame(sz) %>%
    rownames_to_column() %>%
    pivot_longer(names_to = "files",
                 values_to = "size",
                 -rowname) %>%
    rename(backend = rowname) %>%
    mutate(files = as.numeric(sub("X", "", files))) %>%
    ggplot(aes(x = files, y = size, colour = backend)) +
    geom_point() +
    geom_line() +
    scale_x_continuous(breaks=c(1, 5, 10)) +
    ylab("Object size [MB]") +
    xlab("Number of files") +
    theme(legend.position = "none")
@

@@

\begin{figure}[p]
  \centering
<<plot_bench>>=
(p_read + p_sz + p_filt) / p_access +
    plot_layout(heights = c(1, 0.75)) +
    plot_annotation(tag_levels = 'a') &
    theme(axis.title = element_text(size = 10))
@
\caption{(a) Reading time (in seconds) and (b) data size in memory (in
  MB) to read/store 1, 5 and 10 files containing 1431 MS1 (on-disk
  only) and 6103 MS2 (on-disk and in-memory) spectra. (c) Filtering
  benchmark assessed over 10 interactions on in-memory and on-disk
  data containing 6103 MS2 spectra.  (d) Access time to spectra for
  the in-memory (left) and on-disk (right) backends for 1, 10, 100
  1000, 5000 and all 6103 spectra. Benchmarks were performed on a Dell
  XPS laptop with an Intel i5-8250U processor 1.60 GHz (4 cores, 8
  threads), 7.5 GB RAM running Ubuntu 18.04.4 LTS 64-bit
  \textcolor{red}{and an SSD drive}. \textcolor{red}{The data used for
    the benchmarking are a TMT 4-plex experiment acquired on a LTQ
    Orbitrap Velos ((Thermo Fisher Scientific) available in the
    \texttt{msdata} package and described in \citep{Gatto:2014}.}}
\label{fig:bench}
\end{figure}

>>>>

<<time_sz, message=FALSE>>=
load("bench_time_sz.rda")
load("bench_time_2.rda")

sz <- sapply(time_sz, function(x) x[, "sz"])
colnames(sz) <- c(1, 5, 10)
rownames(sz) <- c("in-memory", "on-disk")
sz <- round(sz/(1014^2), 2)

set.seed(1L) ## set jittering
p_read <- 
    time_2 %>%
    magrittr::set_colnames(c("n", "in-memory", "on-disk")) %>%
    pivot_longer(names_to = "backend",
                 values_to = "time", 
                 cols = -n) %>%
    ggplot(aes(x = factor(n), y = time, colour = backend)) + 
    geom_jitter(width = 0.05, size = 2, alpha = 0.5) + 
    geom_smooth(aes(group = backend),
                method = "loess", 
                se = FALSE,
                size = 0.8) + 
    ylab("Reading time [s]") +
    xlab("Number of files") +
    theme(legend.position = c(0.35, 0.86),
          legend.title = element_blank(),
          legend.text=element_text(size = 6))

p_sz <-
    data.frame(sz) %>%
    rownames_to_column() %>%
    pivot_longer(names_to = "files",
                 values_to = "size",
                 -rowname) %>%
    rename(backend = rowname) %>%
    mutate(files = as.numeric(sub("X", "", files))) %>%
    ggplot(aes(x = files, y = size, colour = backend)) +
    geom_point(size = 2, alpha = 0.5) +
    geom_line(size = 0.8) +
    scale_x_continuous(breaks=c(1, 5, 10)) +
    ylab("Object size [MB]") +
    xlab("Number of files") +
    theme(legend.position = "none")
@

@@

\begin{figure}[p]
  \centering
<<plot_bench, warning=FALSE, message=FALSE>>=
(p_read + p_sz + p_filt) / p_access +
    plot_layout(heights = c(1, 0.75)) +
    plot_annotation(tag_levels = 'a') &
    theme(axis.title = element_text(size = 10))
@
\caption{(a) Reading time (triplicates, in seconds) and (b) data size
  in memory (in MB) to read/store 1, 5 and 10 files containing 1431
  MS1 (on-disk only) and 6103 MS2 (on-disk and in-memory) spectra. (c)
  Filtering benchmark assessed over 10 interactions on in-memory and
  on-disk data containing 6103 MS2 spectra.  (d) Access time to
  spectra for the in-memory (left) and on-disk (right) backends for 1,
  10, 100 1000, 5000 and all 6103 spectra. Benchmarks were performed
  on a Dell XPS laptop with an Intel i5-8250U processor 1.60 GHz (4
  cores, 8 threads), 7.5 GB RAM running Ubuntu 18.04.4 LTS 64-bit
  \textcolor{red}{and an SSD drive}. \textcolor{red}{The data used for
    the benchmarking are a TMT 4-plex experiment acquired on a LTQ
    Orbitrap Velos ((Thermo Fisher Scientific) available in the
    \texttt{msdata} package and described in \citep{Gatto:2014}.}}
\label{fig:bench}
\end{figure}

### 2020-07-24 msnbase2.tex
triplicated reading timings

\end{knitrout}
\caption{(a) Reading time (in seconds) and (b) data size in memory (in
  MB) to read/store 1, 5 and 10 files containing 1431 MS1 (on-disk
  only) and 6103 MS2 (on-disk and in-memory) spectra. (c) Filtering
  benchmark assessed over 10 interactions on in-memory and on-disk
  data containing 6103 MS2 spectra.  (d) Access time to spectra for
  the in-memory (left) and on-disk (right) backends for 1, 10, 100
  1000, 5000 and all 6103 spectra. Benchmarks were performed on a Dell
  XPS laptop with an Intel i5-8250U processor 1.60 GHz (4 cores, 8
  threads), 7.5 GB RAM running Ubuntu 18.04.4 LTS 64-bit
  \textcolor{red}{and an SSD drive}. \textcolor{red}{The data used for
    the benchmarking are a TMT 4-plex experiment acquired on a LTQ
    Orbitrap Velos ((Thermo Fisher Scientific) available in the
    \texttt{msdata} package and described in \citep{Gatto:2014}.}}
\label{fig:bench}
\end{figure}

>>>>

\end{knitrout}
\caption{(a) Reading time (triplicates, in seconds) and (b) data size
  in memory (in MB) to read/store 1, 5 and 10 files containing 1431
  MS1 (on-disk only) and 6103 MS2 (on-disk and in-memory) spectra. (c)
  Filtering benchmark assessed over 10 interactions on in-memory and
  on-disk data containing 6103 MS2 spectra.  (d) Access time to
  spectra for the in-memory (left) and on-disk (right) backends for 1,
  10, 100 1000, 5000 and all 6103 spectra. Benchmarks were performed
  on a Dell XPS laptop with an Intel i5-8250U processor 1.60 GHz (4
  cores, 8 threads), 7.5 GB RAM running Ubuntu 18.04.4 LTS 64-bit
  \textcolor{red}{and an SSD drive}. \textcolor{red}{The data used for
    the benchmarking are a TMT 4-plex experiment acquired on a LTQ
    Orbitrap Velos ((Thermo Fisher Scientific) available in the
    \texttt{msdata} package and described in \citep{Gatto:2014}.}}
\label{fig:bench}
\end{figure}
